Mortar-based entropy stable discontinuous Galerkin methods on non-conforming quadrilateral and hexahedral meshes

### 2020-02-27 docs/ESDG_mortar/main.tex
deleting a bunch of stuff and staging old changes

\usepackage{amsmath, amsfonts, amssymb}
\usepackage{stmaryrd} % for llbracket

@@

In this work, we extend entropy stable Gauss collocation schemes to non-conforming meshes.  Entropy stable DG schemes require computing entropy conservative numerical fluxes between volume and surface quadrature nodes.  On conforming tensor product meshes where volume and surface nodes are aligned, flux evaluations are required only between ``lines'' of nodes.  However, on non-conforming meshes, volume and surface nodes are no longer aligned, resulting in a larger number of flux evaluations.  We reduce this expense by introducing an entropy stable mortar-based treatment of non-conforming interfaces via a face-local correction term, and provide truncation error estimates for the resulting discretizations.  Numerical experiments confirm the stability and accuracy of this approach, and we conclude by discussing how to extend mortar treatments of non-conforming interfaces to general elements and arbitrary quadrature rules \cite{chan2017discretely}.  
\end{abstract}

@@

The paper is organized as follows: Section~\ref{sec:0} briefly reviews the derivation of an entropy inequality for a system of nonlinear conservation laws.  Section~\ref{sec:1} introduces ``hybridized'' SBP operators as a unified way to treat both Lobatto and Gauss collocation schemes on conforming meshes of tensor product elements.  Section~\ref{sec:2} describes a naive extension to non-conforming meshes and illustrates why this formulation results in an increase in computational costs.  Section~\ref{sec:3} introduces a mortar-based formulation which addresses such costs, and characterizes the accuracy of the resulting formulations.  Section~\ref{sec:4} describes the extension to curved meshes, as well as more general elements and $p$ non-conformity.  We conclude with numerical validation of theoretical results in Section~\ref{sec:5}.

@@

It is possible to construct energy stable high order discretizations of linear hyperbolic systems using GSBP operators based on Gauss nodes \cite{fernandez2014generalized}.  However, for nonlinear conservation laws, the presence of the dense $\bm{E}$ matrix in the GSBP property (\ref{eq:gsbp}) complicates the imposition of boundary conditions and computation of inter-element numerical fluxes \cite{crean2018entropy, chan2017discretely, chan2018efficient}.  This can be avoided by using ``hybridized'' (or decoupled) SBP operators \cite{chan2017discretely, chenreview, chan2019skew}, which are defined as the block matrix 
\[
\bm{Q}_h = \frac{1}{2}\begin{bmatrix}
\bm{Q}-\bm{Q}^T & \bm{E}^T\bm{B}\\
-\bm{B}\bm{E} & \bm{B}
\end{bmatrix}.
\]
The hybridized SBP operator satisfies a block form of the SBP property
\begin{equation}
\bm{Q}_h + \bm{Q}_h^T = \bm{B}_h, \qquad \bm{B}_h =  \begin{bmatrix}
\bm{0} & \\
& \bm{B}\end{bmatrix}.
\label{eq:hsbp}
\end{equation}
where $\bm{E}$ does not appear in the block boundary matrix on the right hand side.  Here, we have used $\bm{0}, \bm{1}$ to denote a matrix or vector of zeros or ones, where the size is inferred from context.  Note that the matrix $\bm{Q}_h$ also satisfies
\begin{equation}
\bm{Q}_h\bm{1} = 
\frac{1}{2}\begin{bmatrix}
\bm{Q}\bm{1}-\bm{Q}^T\bm{1} + \bm{E}^T\bm{B}\bm{1}\\
-\bm{B}\bm{E}\bm{1} - \bm{B}\bm{1}
\end{bmatrix}= 
\frac{1}{2}\begin{bmatrix}
\LRp{-\bm{Q}^T + \bm{E}^T\bm{B}}\bm{1}\\
\bm{0}
\end{bmatrix}=
\frac{1}{2}\begin{bmatrix}
\bm{Q}\bm{1}\\
\bm{0}
\end{bmatrix} = \bm{0}
\label{eq:Qh1}
\end{equation}
where we have used that $\bm{E}\bm{1} = \bm{1}$ (since $\bm{E}$ is a high order accurate boundary interpolation matrix), $\bm{Q}\bm{1} = \bm{0}$ (since $\bm{Q}$ is a differentiation matrix), and the GSBP property $-\bm{Q}^T + \bm{E}^T\bm{B}\bm{E} = \bm{Q}$

@@

The 2D differentiation and interpolation matrices $\hat{\bm{Q}}_i, \bm{E}$ can now be used to construct 2D hybridized SBP operators.  Let $\hat{\bm{Q}}_{i,h}$ denote the hybridized SBP operator for the $i$th coordinate on the reference element, where $\hat{\bm{Q}}_{i,h}$ is defined as
\begin{align}
\hat{\bm{Q}}_{i,h} = \frac{1}{2}\begin{bmatrix}
\hat{\bm{Q}}_i - \hat{\bm{Q}}_i^T & \bm{E}^T\hat{\bm{B}}_i\\
-\hat{\bm{B}}_i\bm{E} & \hat{\bm{B}}_i
\end{bmatrix}.
\label{eq:hsbpdef}
\end{align}
These matrices satisfy the following multi-dimensional properties:
\begin{lemma}
\label{lemma:hsbprefprops}
Let $\hat{\bm{Q}}_{i}$ be defined as in (\ref{eq:kron}) and let $\hat{\bm{Q}}_{i,h}$ be defined as in (\ref{eq:hsbpdef}).  Then, 
\begin{align}
\hat{\bm{Q}}_{i,h} + \hat{\bm{Q}}_{i,h}^T = \begin{bmatrix}
\bm{0} & \\
& \hat{\bm{B}}_i
\end{bmatrix}, \qquad \hat{\bm{Q}}_{i,h}\bm{1} = \bm{0}.
\end{align}
\end{lemma}
The proof is contained in \cite{chan2018efficient}.  The SBP property follows from the construction of $\hat{\bm{Q}}_{i,h}$, and $\hat{\bm{Q}}_{i,h}\bm{1} = \bm{0}$ is a consequence of the Kronecker product and the same arguments used to derive (\ref{eq:Qh1}).  The construction of 3D differentiation and interpolation matrices proceeds similarly, and the matrices satisfy the same properties as described in Lemma~\ref{lemma:hsbprefprops}.

@@

%Finally, we can extend this formulation to the multi-element case.  
We now construct an entropy conservative formulation on an unstructured curved mesh.  
Suppose the domain $\Omega$ is decomposed into non-overlapping elements $D^k$ which are images of the reference mapping, such that $D^k$ is a differentiable mapping of the reference element $\hat{D} = [-1,1]^d$.  Let $\bm{x}_i$ denote the $i$th physical coordinate on $D^k$ and let $\hat{\bm{x}}_j$ denote the $j$th reference coordinate.  The mapping between reference and physical element induces scaled geometric terms 
\[
G_{ij} = J\pd{\hat{\bm{x}}_j}{\bm{x}_i},
\]
where $J$ denotes the determinant of the Jacobian of the physical-to-reference mapping.  We refer to $J$ as the Jacobian from here onwards.
These geometric terms also relate the normal vectors on reference and physical elements.  Let $\hat{\bm{n}}\hat{J}_f$ denote the outward normal vector on $\hat{D}$, scaled by the reference face Jacobians.  Then, the (scaled) physical normal vectors on $D^k$ are related to $\hat{\bm{n}}\hat{J}_f$ through
\begin{equation}
{n}_iJ_f = \sum_{j=1}^d{G}_{ij}\hat{{n}}_j\hat{J}_f.
\label{eq:nJ_Gnhat}
\end{equation}
We assume that the mesh is watertight, such that the scaled normal vectors are equal and opposite across shared faces between two elements.  We also assume that the geometric terms ${G}_{ij}$ are constructed such that they are polynomials of degree less than or equal to $N$ and satisfy a discrete geometric conservation law (GCL)
\begin{equation}
\sum_{j=1}^d \pd{}{\hat{x}_j} G_{ij} = 0.
\label{eq:dgcl}
\end{equation}
The discrete GCL is satisfied automatically for isoparametric mappings in 2D, and there exist several techniques to enforce the satisfaction of a discrete GCL on various three-dimensional domains \cite{thomas1979geometric, kopriva2006metric, crean2018entropy, chan2018discretely, kozdon2018energy, kopriva2019free}.

@@

Because the reference matrices $\hat{\bm{Q}}_{j,h}$ satisfy the reference SBP properties (\ref{eq:hsbp}), one can show \cite{chan2018discretely} that the physical matrices $\bm{Q}_{i,h}$ satisfy analogous SBP properties.  
\begin{lemma}
\label{lemma:Qhprops}
Suppose the geometric terms $\bm{G}_{ij}$ satisfy the discrete GCL (\ref{eq:dgcl}), the normals are constructed via (\ref{eq:nJ_Gnhat}), and that $\bm{Q}_{i,h}$ is constructed using (\ref{eq:curvedQ}).  Then, 
\[
\bm{Q}_{i,h} + \bm{Q}_{i,h}^T = \begin{bmatrix}
\bm{0} &\\
& \bm{B}_i \end{bmatrix}, \qquad \bm{Q}_{i,h}\bm{1} = \bm{0},
\]
where $\bm{n}_i$ is a vector whose entries are values of the $i$th component of the scaled normals $n_iJ_f$ at face points, $\bm{w}_f$ is the vector of face quadrature weights $w_i$, and $\bm{B}_i = \diag{\bm{n}_i \circ \bm{w}_f}$.
\end{lemma}
\begin{proof}
The proof is identical to that of \cite{chan2018discretely}.  
We review it here for completeness.  Expanding out $\bm{Q}_{i,h} + \bm{Q}_{i,h}^T$ yields
\[
\frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h} + \hat{\bm{Q}}_{j,h}^T\diag{\bm{G}_{ij}} + \hat{\bm{Q}}_{j,h} \diag{\bm{G}_{ij}} +  \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h}^T.
\]
For each term in the sum, we apply the SBP property of $\hat{\bm{Q}}_{j,h}$.  The volume terms cancel, leaving only surface terms
\[
\frac{1}{2}\sum_{j=1}^d \LRp{\hat{\bm{B}}_{j,h}\diag{\bm{G}_{ij}} +  \diag{\bm{G}_{ij}}\hat{\bm{B}}_{j,h}} = \begin{bmatrix}
\bm{0} &\\
& \bm{B}_i \end{bmatrix}.
\]
where we have used (\ref{eq:nJ_Gnhat}) and that $\hat{\bm{B}}_{j,h}$ and $\diag{\bm{G}_{ij}}$ are diagonal matrices.

To show $\bm{Q}_{i,h}\bm{1} = \bm{0}$, we note that
\[
\bm{Q}_{i,h}\bm{1} = \frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h}\bm{1} + \hat{\bm{Q}}_{j,h} \diag{\bm{G}_{ij}}\bm{1} = \frac{1}{2}\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{G}_{ij}
\]
since $\hat{\bm{Q}}_{j,h}\bm{1} = \bm{0}$ by Lemma~\ref{lemma:hsbprefprops}.  Note that, since $\bm{G}_{ij}$ is constructed using polynomial interpolation, $\bm{G}_{ij} = \LRs{G_{ij}(\hat{\bm{x}}), \quad \bm{E}G_{ij}(\hat{\bm{x}})}^T$, where $G_{ij}(\hat{\bm{x}})$ denotes the values of $G_{ij}$ at volume points.  The remaining term can be expanded out
\begin{align*}
\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{G}_{ij} &= \sum_{j=1}^d \begin{bmatrix} \hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}}) - \hat{\bm{Q}}_j^TG_{ij}(\hat{\bm{x}}) + \bm{E}^T\bm{B}\bm{E}G_{ij}(\hat{\bm{x}})\\
\bm{B}\bm{E}G_{ij}(\hat{\bm{x}}) - \bm{B}\bm{E}G_{ij}(\hat{\bm{x}})
\end{bmatrix} \\
&=
\sum_{j=1}^d \begin{bmatrix} \LRp{- \hat{\bm{Q}}_j^T + \bm{E}^T\bm{B}\bm{E}}G_{ij}(\hat{\bm{x}})\\
\bm{0}
\end{bmatrix} = 
\sum_{j=1}^d \begin{bmatrix} \hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}})\\
\bm{0}
\end{bmatrix}  = \bm{0},
\end{align*}
where we have used  the generalized SBP property and that $\sum_{j=1}^d\hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}}) = \bm{0}$, since $G_{ij}$ satisfies the discrete GCL (\ref{eq:dgcl}), $G_{ij}(\hat{\bm{x}})$ is a polynomial of degree less than or equal to $N$, and $\hat{\bm{Q}}_j$ exactly differentiates polynomials of degree $N$.  
\qed\end{proof}

\subsection{Entropy conservative formulations on conforming meshes of mapped elements}

The aforementioned matrices can now be used to construct high order accurate entropy stable and entropy conservative schemes.  We first introduce an entropy conservative numerical flux \cite{tadmor1987numerical}.  Let $\bm{u}_L, \bm{u}_R$ denote left and right states in the conservative variables.  Then, an entropy conservative numerical flux is a vector-valued function $\bm{f}_{i,S}$ for $i = 1,\ldots,d$ which satisfies the following three properties
\begin{gather*}
\bm{f}_{i,S}\LRp{\bm{u},\bm{u}} = \bm{f}_i(\bm{u}), \qquad \text{(consistency)}\\
\bm{f}_{i,S}\LRp{\bm{u}_L,\bm{u}_R} = \bm{f}_{i,S}(\bm{u}_R,\bm{u}_L), \qquad \text{(symmetry)}\\
\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_{i,S}\LRp{\bm{u}_L,\bm{u}_R} = \psi_i(\bm{u}_L) - \psi_i(\bm{u}_R), \qquad \text{(conservation)}.
\end{gather*}
Here, $\bm{v}_L, \bm{v}_R$ are the entropy variables evaluated at $\bm{u}_L, \bm{u}_R$, and $\psi_i$ is the $i$th entropy potential which appears in (\ref{eq:weakentropyineq}).  
Then, an entropy conservative scheme on a mapped element is given by
\begin{gather}
{\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
\sum_{i=1}^d \LRp{2{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T{\bm{B}}_i\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_f)} = 0 \label{eq:esdg}\\
\LRp{\bm{F}_i}_{jk} = \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}, \qquad \bm{f}_i^* = \bm{f}_{i,S}\LRp{\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f} \nonumber
\end{gather}
Here, the exterior state $\tilde{\bm{u}}_f^+$ used to evaluate the numerical flux $\bm{f}_i^*$ corresponds either to interface values on a neighboring element or an exterior state used to enforce boundary conditions, and $\circ$ denotes the matrix Hadamard product.

%\note{Mention proof}
%Then, an entropy conservative scheme on the reference interval is given as
%\begin{gather}
%\bm{M}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
%\LRp{2\bm{Q}_h \circ \bm{F}}\bm{1} + \bm{E}^T\bm{B}\LRp{\bm{f}^*-\bm{f}(\tilde{\bm{u}}_f)} = 0 \label{eq:esdg1D}\\ 
%\tilde{\bm{u}} = \begin{bmatrix}
%\bm{u}_h\\
%\tilde{\bm{u}}_f
%\end{bmatrix}, \qquad \tilde{\bm{u}}_f = \bm{u}\LRp{\bm{E}\bm{v}(\bm{u}_h)},\nonumber \\ 
%\LRp{\bm{F}}_{jk} = \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}, \qquad \bm{f}^* = \bm{f}_S\LRp{\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f}.\nonumber
%\end{gather}
%Here, $\tilde{\bm{u}}_f^+$ denotes some exterior state which can be used to enforce boundary conditions or coupling between different elements

>>>>

\usepackage{stmaryrd} % for llbracket
\usepackage{amsmath, amsfonts, amssymb}

@@

In this work, we extend entropy stable Gauss collocation schemes to non-conforming meshes.  Entropy stable DG schemes require computing entropy conservative numerical fluxes between volume and surface quadrature nodes.  On conforming tensor product meshes where volume and surface nodes are aligned, flux evaluations are required only between ``lines'' of nodes.  However, on non-conforming meshes, volume and surface nodes are no longer aligned, resulting in a larger number of flux evaluations.  We reduce this expense by introducing an entropy stable mortar-based treatment of non-conforming interfaces via a face-local correction term, and provide necessary conditions for high order accuracy.  Numerical experiments confirm the stability and accuracy of this approach.  
\end{abstract}

@@

The paper is organized as follows: Section~\ref{sec:0} briefly reviews the derivation of an entropy inequality for a system of nonlinear conservation laws.  Section~\ref{sec:1} introduces ``hybridized'' SBP operators as a unified way to treat both Lobatto and Gauss collocation schemes on conforming meshes of tensor product elements.  Section~\ref{sec:2} describes a naive extension to non-conforming meshes and illustrates why this formulation results in an increase in computational costs.  Section~\ref{sec:3} introduces a mortar-based formulation on curved meshes which addresses such costs, and characterizes the accuracy of the resulting formulations.  We conclude with numerical validation of theoretical results in Section~\ref{sec:4}.

@@

It is possible to construct energy stable high order discretizations of linear hyperbolic systems using GSBP operators based on Gauss nodes \cite{fernandez2014generalized}.  However, for nonlinear conservation laws, the presence of the dense $\bm{E}$ matrix in the GSBP property (\ref{eq:gsbp}) complicates the imposition of boundary conditions and computation of inter-element numerical fluxes \cite{crean2018entropy, chan2017discretely, chan2018efficient}.  This can be avoided by using ``hybridized'' (also referred to as decoupled) SBP operators \cite{chan2017discretely, chenreview, chan2019skew}, which are defined as the block matrix 
\[
\bm{Q}_h = \frac{1}{2}\begin{bmatrix}
\bm{Q}-\bm{Q}^T & \bm{E}^T\bm{B}\\
-\bm{B}\bm{E} & \bm{B}
\end{bmatrix}.
\]
The hybridized SBP operator satisfies a block form of the SBP property
\begin{equation}
\bm{Q}_h + \bm{Q}_h^T = \bm{B}_h, \qquad \bm{B}_h =  \begin{bmatrix}
\bm{0} & \\
& \bm{B}\end{bmatrix}.
\label{eq:hsbp}
\end{equation}
where $\bm{E}$ does not appear in the block boundary matrix on the right hand side.  Here, we have used $\bm{0}, \bm{1}$ to denote a matrix or vector of zeros or ones, where the size is inferred from context.  Note that the matrix $\bm{Q}_h$ also satisfies
\begin{equation}
\bm{Q}_h\bm{1} = 
\frac{1}{2}\begin{bmatrix}
\bm{Q}\bm{1}-\bm{Q}^T\bm{1} + \bm{E}^T\bm{B}\bm{1}\\
-\bm{B}\bm{E}\bm{1} - \bm{B}\bm{1}
\end{bmatrix}= 
\frac{1}{2}\begin{bmatrix}
\LRp{-\bm{Q}^T + \bm{E}^T\bm{B}}\bm{1}\\
\bm{0}
\end{bmatrix}=
\frac{1}{2}\begin{bmatrix}
\bm{Q}\bm{1}\\
\bm{0}
\end{bmatrix} = \bm{0}
\label{eq:Qh1}
\end{equation}
where we have used that $\bm{E}\bm{1} = \bm{1}$ (since $\bm{E}$ is a high order accurate boundary interpolation matrix), $\bm{Q}\bm{1} = \bm{0}$ (since $\bm{Q}$ is a differentiation matrix), and the GSBP property $-\bm{Q}^T + \bm{E}^T\bm{B}\bm{E} = \bm{Q}$

@@

The 2D differentiation and interpolation matrices $\hat{\bm{Q}}_i, \bm{E}$ can now be used to construct 2D hybridized SBP operators.  Let $\hat{\bm{Q}}_{i,h}$ denote the hybridized SBP operator for the $i$th coordinate on the reference element, where $\hat{\bm{Q}}_{i,h}$ is defined as
\begin{align}
\hat{\bm{Q}}_{i,h} = \frac{1}{2}\begin{bmatrix}
\hat{\bm{Q}}_i - \hat{\bm{Q}}_i^T & \bm{E}^T\hat{\bm{B}}_i\\
-\hat{\bm{B}}_i\bm{E} & \hat{\bm{B}}_i
\end{bmatrix}.
\label{eq:hsbpdef}
\end{align}
These matrices satisfy the following multi-dimensional properties:
\begin{lemma}
\label{lemma:hsbprefprops}
Let $\hat{\bm{Q}}_{i}$ be defined as in (\ref{eq:kron}) and let $\hat{\bm{Q}}_{i,h}$ be defined as in (\ref{eq:hsbpdef}).  Then, 
\begin{align}
\hat{\bm{Q}}_{i,h} + \hat{\bm{Q}}_{i,h}^T = \begin{bmatrix}
\bm{0} & \\
& \hat{\bm{B}}_i
\end{bmatrix}, \qquad \hat{\bm{Q}}_{i,h}\bm{1} = \bm{0}.
\end{align}
\end{lemma}
The proof is found in \cite{chan2018efficient}.  The SBP property follows from the construction of $\hat{\bm{Q}}_{i,h}$, and $\hat{\bm{Q}}_{i,h}\bm{1} = \bm{0}$ is a consequence of properties of the Kronecker product and the same arguments used to derive (\ref{eq:Qh1}).  The construction of 3D differentiation and interpolation matrices proceeds similarly, and the matrices satisfy the same properties as described in Lemma~\ref{lemma:hsbprefprops}.

@@

%Finally, we can extend this formulation to the multi-element case.  
We now construct an entropy conservative formulation on an unstructured curved mesh.  
Suppose the domain $\Omega$ is decomposed into non-overlapping elements $D^k$ which are images of the reference mapping, such that $D^k$ is a differentiable mapping of the reference element $\hat{D} = [-1,1]^d$.  Let $\bm{x}_i$ denote the $i$th physical coordinate on $D^k$ and let $\hat{\bm{x}}_j$ denote the $j$th reference coordinate.  The mapping between reference and physical element induces scaled geometric terms 
\[
G_{ij} = J\pd{\hat{\bm{x}}_j}{\bm{x}_i},
\]
where $J$ denotes the determinant of the Jacobian of the physical-to-reference mapping.  We refer to $J$ as the Jacobian from here onwards.
These geometric terms also relate the normal vectors on reference and physical elements.  Let $\hat{\bm{n}}\hat{J}_f$ denote the outward normal vector on $\hat{D}$, scaled by the reference face Jacobians.  Then, the (scaled) physical normal vectors on $D^k$ are related to $\hat{\bm{n}}\hat{J}_f$ through
\begin{equation}
{n}_iJ_f = \sum_{j=1}^d{G}_{ij}\hat{{n}}_j\hat{J}_f.
\label{eq:nJ_Gnhat}
\end{equation}
We assume that the mesh is watertight, such that the scaled normal vectors are equal and opposite across shared faces between two elements.  We also assume that the geometric terms ${G}_{ij}$ are constructed such that they are polynomials of degree less than or equal to $N$ and satisfy a discrete geometric conservation law (GCL)
\begin{equation}
\sum_{j=1}^d \pd{}{\hat{x}_j} G_{ij} = 0.
\label{eq:dgcl}
\end{equation}
The discrete GCL is satisfied automatically for isoparametric mappings in 2D, and there exist several techniques to enforce the satisfaction of a discrete GCL on various three-dimensional domains \cite{thomas1979geometric, kopriva2006metric, crean2018entropy, chan2018discretely, kozdon2018energy, kopriva2019free}.  We will discuss specific approaches for non-conforming meshes in Section~\ref{sec:mapped} and Appendix~\ref{app:A}.

@@

Because the reference matrices $\hat{\bm{Q}}_{j,h}$ satisfy the reference SBP properties (\ref{eq:hsbp}), one can show \cite{chan2018discretely} that the physical matrices $\bm{Q}_{i,h}$ satisfy analogous SBP properties.  
\begin{lemma}
\label{lemma:Qhprops}
Suppose the geometric terms $\bm{G}_{ij}$ satisfy the discrete GCL (\ref{eq:dgcl}), the normals are constructed via (\ref{eq:nJ_Gnhat}), and that $\bm{Q}_{i,h}$ is constructed using (\ref{eq:curvedQ}).  Then, 
\[
\bm{Q}_{i,h} + \bm{Q}_{i,h}^T = \begin{bmatrix}
\bm{0} &\\
& \bm{B}_i \end{bmatrix}, \qquad \bm{Q}_{i,h}\bm{1} = \bm{0},
\]
where $\bm{n}_i$ is a vector whose entries are values of the $i$th component of the scaled normals $n_iJ_f$ at face points, $\bm{w}_f$ is the vector of face quadrature weights $w_i$, and $\bm{B}_i = \diag{\bm{n}_i \circ \bm{w}_f}$.
\end{lemma}
The proof is given in \cite{chan2018discretely}.  
%\begin{proof}
%The proof is identical to that of \cite{chan2018discretely}.  
%We review it here for completeness.  Expanding out $\bm{Q}_{i,h} + \bm{Q}_{i,h}^T$ yields
%\[
%\frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h} + \hat{\bm{Q}}_{j,h}^T\diag{\bm{G}_{ij}} + \hat{\bm{Q}}_{j,h} \diag{\bm{G}_{ij}} +  \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h}^T.
%\]
%For each term in the sum, we apply the SBP property of $\hat{\bm{Q}}_{j,h}$.  The volume terms cancel, leaving only surface terms
%\[
%\frac{1}{2}\sum_{j=1}^d \LRp{\hat{\bm{B}}_{j,h}\diag{\bm{G}_{ij}} +  \diag{\bm{G}_{ij}}\hat{\bm{B}}_{j,h}} = \begin{bmatrix}
%\bm{0} &\\
%& \bm{B}_i \end{bmatrix}.
%\]
%where we have used (\ref{eq:nJ_Gnhat}) and that $\hat{\bm{B}}_{j,h}$ and $\diag{\bm{G}_{ij}}$ are diagonal matrices.  
%
%To show $\bm{Q}_{i,h}\bm{1} = \bm{0}$, we note that
%\[
%\bm{Q}_{i,h}\bm{1} = \frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h}\bm{1} + \hat{\bm{Q}}_{j,h} \diag{\bm{G}_{ij}}\bm{1} = \frac{1}{2}\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{G}_{ij}
%\]
%since $\hat{\bm{Q}}_{j,h}\bm{1} = \bm{0}$ by Lemma~\ref{lemma:hsbprefprops}.  Note that, since $\bm{G}_{ij}$ is constructed using polynomial interpolation, $\bm{G}_{ij} = \LRs{G_{ij}(\hat{\bm{x}}), \quad \bm{E}G_{ij}(\hat{\bm{x}})}^T$, where $G_{ij}(\hat{\bm{x}})$ denotes the values of $G_{ij}$ at volume points.  The remaining term can be expanded out
%\begin{align*}
%\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{G}_{ij} &= \sum_{j=1}^d \begin{bmatrix} \hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}}) - \hat{\bm{Q}}_j^TG_{ij}(\hat{\bm{x}}) + \bm{E}^T\bm{B}\bm{E}G_{ij}(\hat{\bm{x}})\\
%\bm{B}\bm{E}G_{ij}(\hat{\bm{x}}) - \bm{B}\bm{E}G_{ij}(\hat{\bm{x}})
%\end{bmatrix} \\
%&=
%\sum_{j=1}^d \begin{bmatrix} \LRp{- \hat{\bm{Q}}_j^T + \bm{E}^T\bm{B}\bm{E}}G_{ij}(\hat{\bm{x}})\\
%\bm{0}
%\end{bmatrix} = 
%\sum_{j=1}^d \begin{bmatrix} \hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}})\\
%\bm{0}
%\end{bmatrix}  = \bm{0},
%\end{align*}
%where we have used  the generalized SBP property and that $\sum_{j=1}^d\hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}}) = \bm{0}$, since $G_{ij}$ satisfies the discrete GCL (\ref{eq:dgcl}), $G_{ij}(\hat{\bm{x}})$ is a polynomial of degree less than or equal to $N$, and $\hat{\bm{Q}}_j$ exactly differentiates polynomials of degree $N$.  
%\qed\end{proof}

\subsection{Entropy conservative formulations on conforming meshes of mapped elements}
\label{sec:mapped}

The aforementioned matrices can now be used to construct high order accurate entropy stable and entropy conservative schemes.  We first introduce an entropy conservative numerical flux \cite{tadmor1987numerical}.  Let $\bm{u}_L, \bm{u}_R$ denote left and right states in the conservative variables.  Then, an entropy conservative numerical flux is a vector-valued function $\bm{f}_{i,S}$ for $i = 1,\ldots,d$ which satisfies the following three properties
\begin{gather*}
\bm{f}_{i,S}\LRp{\bm{u},\bm{u}} = \bm{f}_i(\bm{u}), \qquad \text{(consistency)}\\
\bm{f}_{i,S}\LRp{\bm{u}_L,\bm{u}_R} = \bm{f}_{i,S}(\bm{u}_R,\bm{u}_L), \qquad \text{(symmetry)}\\
\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_{i,S}\LRp{\bm{u}_L,\bm{u}_R} = \psi_i(\bm{u}_L) - \psi_i(\bm{u}_R), \qquad \text{(conservation)}.
\end{gather*}
Here, $\bm{v}_L, \bm{v}_R$ are the entropy variables evaluated at $\bm{u}_L, \bm{u}_R$, and $\psi_i$ is the $i$th entropy potential which appears in (\ref{eq:weakentropyineq}).  
Then, an entropy conservative scheme on a mapped element is given by
{\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
\sum_{i=1}^d \LRp{2{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T{\bm{B}}_i\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_f)} = 0 \label{eq:esdg}\\
\LRp{\bm{F}_i}_{jk} = \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}, \qquad \bm{f}_i^* = \bm{f}_{i,S}\LRp{\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f}, \nonumber
where $\bm{M} = \hat{\bm{M}} \circ \diag{\bm{J}}$ is the diagonal mass matrix scaled by the determinant of the geometric Jacobian $J$ evaluated at nodal points.  
Here, the exterior state $\tilde{\bm{u}}_f^+$ used to evaluate the numerical flux $\bm{f}_i^*$ corresponds either to interface values on a neighboring element or an exterior state used to enforce boundary conditions, and $\circ$ denotes the matrix Hadamard product.  Assuming continuity in time, the formulation (\ref{eq:esdg}) is entropy conservative in the sense that
\bm{1}^T\bm{M}\td{S(\bm{u}_h)}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}_i\LRp{\bm{v}_f^T\bm{f}_i^* - \psi_i(\tilde{\bm{u}}_f)} = 0, \qquad \bm{v}_f =\bm{E}\bm{v}(\bm{u}_h).
The proof uses similar techniques as proofs in other papers \cite{chen2017entropy, crean2018entropy, chan2017discretely, chan2019skew}
%\begin{theorem}
%\label{thm:esdg}
%Assuming continuity in time, the formulation (\ref{eq:esdg}) is entropy conservative in the sense that
%\bm{1}^T\bm{M}\td{S(\bm{u}_h)}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}_i\LRp{\bm{v}_f^T\bm{f}_i^* - \psi_i(\tilde{\bm{u}}_f)} = 0, \qquad \bm{v}_f =\bm{E}\bm{v}(\bm{u}_h).
%\end{theorem}
%\begin{proof}
%The proof uses similar techniques as proofs in other papers \cite{chen2017entropy, crean2018entropy, chan2017discretely, chan2019skew}, and we repeat it for completeness.  We begin by testing (\ref{eq:esdg}) with the vector of entropy variables evaluated at nodal points $\bm{v}(\bm{u}_h)$.  We apply the chain rule in time and use that $\bm{M}$ is diagonal to show that 
%\bm{v}(\bm{u}_h)^T\bm{M}\td{\bm{u}}{t} = \bm{1}^T\bm{M}\LRp{\bm{v}(\bm{u}_h)\circ \td{\bm{u}_h}{t}} =  \bm{1}^T\bm{M}\td{S(\bm{u}_h)}{t}.
%For the volume term, we use the SBP property to arrive at
%\begin{align*}
%&\bm{v}(\bm{u}_h)^T\begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T\LRp{2\bm{Q}_{i,h} \circ \bm{F}_i}\bm{1} = \\
%&\tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_{i,h}-\bm{Q}_{i,h}^T} \circ \bm{F}_i}\bm{1}  + \LRp{\bm{E}\bm{v}(\bm{u}_h)}^T\LRp{\bm{B}_{i,h} \circ\bm{F}_i}\bm{1},
%\end{align*}
%where we have introduced $\tilde{\bm{v}} = \LRp{\begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}\bm{v}(\bm{u}_h)}$.  
%Note that, by the consistency of the entropy conservative flux $\bm{f}_{i,S}$, the diagonal of $\bm{F}_i$ is the vector $\bm{f}_i(\tilde{\bm{u}}_f)$.  Using that $\bm{B}_{i,h}$ is diagonal (with non-zero diagonal sub-block $\bm{B}_i$), the latter term reduces to 
%\LRp{\bm{E}\bm{v}(\bm{u}_h)}^T\LRp{\bm{B}_{i,h} \circ\bm{F}_i}\bm{1} = \bm{v}_f^T \bm{B}_i\bm{f}_i(\tilde{\bm{u}}_f) =  \bm{1}^T \bm{B}_i\LRp{\bm{v}_f^T\bm{f}_i(\tilde{\bm{u}}_f)}.
%The final part of the proof uses that 
%\begin{align*}
